---
title: "Event Study: School Safety Program (March 2026)"
author: "Gabriela Eastwood"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    code-summary: "Show code"
    embed-resources: true
    theme: cosmo
execute:
  warning: false
  message: false
---

## Overview

This document presents two event-study estimates of the causal effect of a
school safety program on municipal violent incidents.

**Design summary**

| Feature | Detail |
|---|---|
| Treatment | School safety program adopted **March 2026** |
| Supplement | Roads component added for half of treated municipalities, **October 2026** |
| Announcement | December 2025 (potential anticipation at k = −1) |
| Outcome | Monthly violent incidents per municipality |
| Identification | Two-way fixed effects (municipality + month); SEs clustered by municipality |
| Reference period | k = −2 (January 2026) |
| Lead binning | All leads ≤ −12 collapsed into one bin (k = −12) |

The reference period is set to **k = −2** rather than k = −1 because the
December 2025 announcement is likely to contaminate the period immediately
before adoption (k = −1) through anticipation effects.

---

## Setup

```{r}
#| label: setup

library(fixest)   # feols(): fixed effects OLS, clustered SEs
library(dplyr)    # data manipulation
library(ggplot2)  # plotting
```

---

## Data

```{r}
#| label: load-data

df <- read.csv("hsif_week6_day2_synthetic.csv")
glimpse(df)
```

### Time index and event time

We convert the `month` string to a proper date, then to a compact integer
time index `t`. Event time `k` measures months relative to the adoption date
(March 2026 ≡ k = 0).

```{r}
#| label: time-index

# Month string → running index t (1, 2, …)
df <- df %>%
  mutate(month_date = as.Date(paste0(month, "-01")),
         t = as.integer(factor(month_date)))

# Event time: k = t − t_adopt; adoption month = March 2026 → k = 0
t_adopt <- df$t[df$month == "2026-03"][1]
df <- df %>% mutate(k = t - t_adopt)

cat("Adoption time index:", t_adopt,
    "\nEvent-time range:", range(df$k), "\n")
```

### Lead binning and reference coding

Distant leads (k ≤ −12) are collapsed into a single bin to keep the
pre-trend plot legible and well-identified. Untreated municipalities are
assigned the reference value (k = −2) so they contribute no event-time
dummies and identify the counterfactual purely through fixed effects.

```{r}
#| label: event-coding

REF <- -2   # reference period

# Bin extreme leads
df <- df %>% mutate(k_bin = pmax(k, -12))

# Pooled event-time variable: untreated → reference, treated → k_bin
df <- df %>% mutate(k_pool = if_else(treated == 1, k_bin, REF))

# Interacted variables: one per treatment arm, both equal REF outside the arm
df <- df %>%
  mutate(k_schools = if_else(treated == 1 & supplement == 0, k_bin, REF),
         k_roads   = if_else(treated == 1 & supplement == 1, k_bin, REF))
```

Quick balance check — observations per treatment arm:

```{r}
#| label: balance-check

df %>%
  distinct(municipality_id, treated, supplement) %>%
  count(treated, supplement, name = "n_municipalities")
```

---

## Figure 1 — Pooled Event Study

**All treated municipalities vs. untreated controls.**

Pre-trends should be flat near zero. A small dip at k = −1 is consistent with
anticipation effects following the December 2025 announcement. Post-adoption
coefficients capture the average effect of the program across both arms.

```{r}
#| label: pooled-estimation

m_pool <- feols(incidents ~ i(k_pool, ref = REF) |
                  municipality_id + t,
                data = df, cluster = ~municipality_id)

summary(m_pool)
```

```{r}
#| label: pooled-plot
#| fig-width: 8
#| fig-height: 4.2
#| fig-cap: "Pooled event study. Dots are OLS estimates; bars are 95% CIs. The open dot at k = −2 marks the reference period (estimate normalized to zero). The dashed vertical line sits between k = −1 and k = 0, marking the onset of treatment."

coefs_pool <- data.frame(
  k   = as.numeric(gsub("k_pool::", "", names(coef(m_pool)))),
  est = coef(m_pool),
  lo  = confint(m_pool)[, 1],
  hi  = confint(m_pool)[, 2]
) %>%
  bind_rows(data.frame(k = REF, est = 0, lo = NA, hi = NA)) %>%
  arrange(k)

ggplot(coefs_pool, aes(x = k, y = est)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
  geom_vline(xintercept = -0.5, linetype = "dashed", color = "gray50") +
  geom_errorbar(aes(ymin = lo, ymax = hi), width = 0.25, color = "gray30") +
  geom_point(color = "gray30", size = 1.8) +
  geom_point(data = filter(coefs_pool, k == REF),
             shape = 21, fill = "white", color = "gray30", size = 1.8) +
  labs(x = "event time k (months since March 2026)",
       y = "effect on violent incidents",
       title = "Pooled event study: all treated vs. untreated (ref k = −2)") +
  theme_minimal(base_size = 11) +
  theme(plot.title = element_text(color = "#800000"))
```

---

## Figure 2 — Interacted Event Study

**Schools-only vs. schools + roads (supplement).**

A single regression with two sets of event-time dummies — one per arm —
measured against the same untreated control group through the shared fixed
effects. The two series should track together through k = 6 (September 2026)
and diverge from k = 7 (October 2026) when the supplement begins. The gap
between them at the panel's end is the supplement's incremental effect.

```{r}
#| label: interacted-estimation

m_int <- feols(incidents ~ i(k_schools, ref = REF) + i(k_roads, ref = REF) |
                 municipality_id + t,
               data = df, cluster = ~municipality_id)

summary(m_int)
```

```{r}
#| label: interacted-plot
#| fig-width: 8
#| fig-height: 4.2
#| fig-cap: "Interacted event study. Gray = schools-only arm; maroon = schools + roads supplement. Estimates are dodged slightly to avoid overlapping CIs. The dotted vertical line marks October 2026, when the roads supplement began."

ci <- confint(m_int)
coefs_int <- data.frame(
  name = names(coef(m_int)),
  est  = coef(m_int),
  lo   = ci[, 1],
  hi   = ci[, 2]
) %>%
  mutate(
    group = if_else(grepl("k_schools", name), "schools only", "schools + roads"),
    k     = as.numeric(gsub("k_schools::|k_roads::", "", name))
  ) %>%
  bind_rows(
    data.frame(k = REF, est = 0, lo = NA, hi = NA,
               group = c("schools only", "schools + roads"))
  )

# Month index of the October 2026 supplement
t_supp_k <- df$t[df$month == "2026-10"][1] - t_adopt

ggplot(coefs_int, aes(x = k, y = est, color = group)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
  geom_vline(xintercept = -0.5, linetype = "dashed", color = "gray50") +
  geom_vline(xintercept = t_supp_k - 0.5, linetype = "dotted",
             color = "#B3541E", linewidth = 0.7) +
  annotate("text", x = t_supp_k - 0.5, y = 2.5,
           label = "Oct 2026: supplement begins",
           color = "#B3541E", size = 3, hjust = 0.5) +
  geom_errorbar(aes(ymin = lo, ymax = hi), width = 0.25,
                position = position_dodge(width = 0.45)) +
  geom_point(size = 1.8, position = position_dodge(width = 0.45)) +
  scale_color_manual(values = c("schools only" = "gray30",
                                "schools + roads" = "#800000")) +
  labs(x = "event time k (months since March 2026)",
       y = "effect on violent incidents",
       title = "Interacted event study: schools-only vs. schools + roads (ref k = −2)",
       color = NULL) +
  theme_minimal(base_size = 11) +
  theme(plot.title  = element_text(color = "#800000"),
        legend.position = c(0.15, 0.15))
```

---

## Regression Table

Both models side-by-side with clustered standard errors.

```{r}
#| label: etable

etable(m_pool, m_int,
       cluster      = ~municipality_id,
       headers      = c("Pooled", "Interacted"),
       title        = "Event-study estimates (outcome: violent incidents)")
```

---

## Interpretation Guide

**Pre-trend check (leads, k < −2)**
: Coefficients should be statistically indistinguishable from zero. A
  systematic upward or downward trend would indicate that treated and control
  municipalities were on different trajectories before the program — a
  violation of parallel trends.

**Anticipation (k = −1)**
: A negative coefficient here is plausible given the December 2025
  announcement. It does *not* invalidate the design but should be noted.

**Impact (k ≥ 0)**
: The pooled post-period coefficients capture the average intention-to-treat
  effect. Negative values indicate a reduction in violent incidents.

**Supplement gap (k ≥ 7)**
: In the interacted plot, any divergence between the two series after October
  2026 is attributable to the roads supplement. The gap's magnitude and its
  CIs determine whether the supplement has a statistically distinguishable
  incremental effect.
