---
title: "Aid Under Fire: Replication Memo"
subtitle: "Crost, Felter & Johnston (2014, AER) — HSIF Week 6 Day 4"
author: "Gabriela Eastwood"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    code-summary: "Show code"
    embed-resources: true
    theme: cosmo
execute:
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false

library(rdrobust)   # MSE-optimal bandwidths, robust bias-corrected inference
library(rddensity)  # density test
library(fixest)     # event study DiD
library(dplyr)
library(ggplot2)
library(haven)

# ---------------------------------------------------------------------
# VARIABLE MAP (filled from codebook exploration)
# ---------------------------------------------------------------------
DATA_FILE  <- "AUF_replication_data.dta"
V_RANK     <- "rank"              # within-province poverty rank (reconstructed)
V_CUT_N    <- "num_muni"         # number of municipalities in province
V_OUTCOME  <- "totcas"           # total conflict casualties
V_CAS_INS  <- "eninit_cas"       # insurgent-initiated casualties
V_CAS_GOV  <- "govinit_cas"      # government-initiated casualties
V_RECEIPT  <- "kalahimunicipality"# actual KALAHI-CIDSS receipt
V_MUNI     <- "municipality"
V_TIME     <- "month"
V_START    <- "aff1"             # program start timing

df <- read_dta(DATA_FILE) %>%
  mutate(
    cutoff_rank = ceiling(.data[[V_CUT_N]] / 4),   # 1st quartile threshold
    r = .data[[V_RANK]] - cutoff_rank,              # centered running variable
    above = as.integer(r >= 0),
    elig  = above
  )
```

---

## Section 1 — Context and the Ranking Problem

*(~250 words | Stage 1: 9:30–9:55 | No estimation)*

**Setting.** Crost, Felter & Johnston (2014) study whether the Philippine
government's KALAHI-CIDSS community development program reduced civil conflict.
KALAHI-CIDSS targeted the poorest municipalities within each province, defined
by a poverty-rank cutoff: municipalities ranked in the bottom quartile of their
province by a poverty index were declared eligible. The authors exploit this
cutoff in a regression discontinuity design, comparing municipalities just
below and just above the eligibility threshold on conflict casualties.

**The ranking problem.** The program's internal targeting list — the actual
poverty rankings used at implementation — is not publicly available. The
replication package reconstructs ranks from published provincial poverty
estimates, but this reconstruction is imperfect. Survey-based poverty estimates
carry measurement error, and the ranking operation amplifies that error near
the cutoff: two municipalities with nearly identical true poverty rates may
swap sides of the cutoff purely due to noise in the underlying estimates.

**Consequences for identification.**

1. *Misclassification near the cutoff.* Municipalities within a rank or two of
   the threshold may be assigned the wrong eligibility status. This attenuates
   the first stage (the jump in actual receipt is smaller than the jump in true
   eligibility) and, under classical measurement error, biases the RD estimate
   toward zero.

2. *Which checks become indirect.* Covariate smoothness tests check whether
   observable pre-determined characteristics are continuous at the reconstructed
   cutoff — but if misclassification is random, covariates may appear smooth
   even when true assignment is noisy. The density test is similarly indirect:
   manipulation of the reconstructed rank is neither necessary nor sufficient
   for manipulation of the true rank.

3. *Which specification targets the problem.* The **donut RD** drops
   municipalities within $|r| \leq 1$ or $|r| \leq 2$ of the reconstructed
   cutoff, restricting estimation to municipalities where the reconstructed and
   true rankings are more likely to agree. If the donut estimate is stable or
   larger than the full-sample estimate, it suggests misclassification was
   attenuating the main result.

---

## Section 2 — Main Results, Modernized

*(~350 words + Figure Array 1 | Stage 2: 9:55–10:40)*

```{r}
#| label: density-check
#| fig-width: 7
#| fig-height: 3.2
#| fig-cap: "Density of the within-province poverty rank centered at the cutoff. No spike below or pile-up above the cutoff is consistent with no manipulation of the reconstructed ranking."

# Collapse to municipality level: sum casualties across all periods,
# keep time-invariant covariates (rank, cutoff, receipt) from first row.
df_muni <- df %>%
  group_by(.data[[V_MUNI]]) %>%
  summarise(
    totcas    = sum(totcas,    na.rm = TRUE),
    eninit_cas = sum(eninit_cas, na.rm = TRUE),
    govinit_cas = sum(govinit_cas, na.rm = TRUE),
    r         = first(r),
    above     = first(above),
    kalahimunicipality = first(kalahimunicipality),
    .groups   = "drop"
  )

ggplot(df_muni, aes(x = r)) +
  geom_histogram(binwidth = 1, fill = "gray70", color = "white") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "#800000") +
  labs(x = "poverty rank (centered at cutoff)", y = "municipalities",
       title = "Density of the running variable") +
  theme_minimal(base_size = 11) +
  theme(plot.title = element_text(color = "#800000"))
```

```{r}
#| label: rd-plot
#| fig-cap: "RD plot: total conflict casualties against centered poverty rank. Bin means shown; local linear fits on each side of the cutoff from the h = 6 window."

rdplot(y = df_muni[[V_OUTCOME]], x = df_muni$r, c = 0,
       x.label = "poverty rank (centered at cutoff)",
       y.label = "conflict casualties",
       title = "")
```

```{r}
#| label: paper-spec

# Replicate paper: h = 6, triangular kernel
cat("=== Paper spec: h = 6, triangular ===\n")
summary(rdrobust(df_muni[[V_OUTCOME]], df_muni$r, c = 0,
                 h = 6, kernel = "triangular", masspoints = "off"))
```

```{r}
#| label: modern-inference

# With only 51 municipalities, MSE-optimal bandwidth selection is unreliable.
# Use fixed, substantively motivated integer bandwidths throughout.
# h = 4: narrow window (ranks -4 to 3), closest to the cutoff.
# h = 8: wider window, trades bias for precision.
cat("=== h = 4 (narrow) ===\n")
summary(rdrobust(df_muni[[V_OUTCOME]], df_muni$r, c = 0, h = 4, masspoints = "off"))

cat("\n=== h = 8 (wide) ===\n")
summary(rdrobust(df_muni[[V_OUTCOME]], df_muni$r, c = 0, h = 8, masspoints = "off"))
```

```{r}
#| label: donut

for (r_hole in c(1, 2)) {
  d <- filter(df_muni, abs(r) > r_hole)
  cat("\n--- Donut r =", r_hole, "(n =", nrow(d), ") ---\n")
  result <- tryCatch(
    summary(rdrobust(d[[V_OUTCOME]], d$r, c = 0, h = 6, masspoints = "off")),
    error = function(e) cat("Note: too few distinct rank values for bias-corrected inference at this donut/bandwidth combination.\n")
  )
  if (!is.null(result)) print(result)
}
```

```{r}
#| label: spec-chart
#| fig-width: 7
#| fig-height: 4
#| fig-cap: "Specification chart: six estimates with 95% CIs on one axis. Bias-corrected Robust rows throughout. Stable estimates across bandwidth and donut specifications support the main finding."

# Collect all six specs into a data frame
collect_rd <- function(label, fit) {
  data.frame(label = label,
             est   = fit$coef["Robust", 1],
             lo    = fit$ci["Robust", 1],
             hi    = fit$ci["Robust", 2])
}

specs <- bind_rows(
  collect_rd("Paper (h = 6)",  rdrobust(df_muni[[V_OUTCOME]], df_muni$r, c = 0, h = 6, kernel = "triangular", masspoints = "off")),
  collect_rd("h = 4 (narrow)", rdrobust(df_muni[[V_OUTCOME]], df_muni$r, c = 0, h = 4, masspoints = "off")),
  collect_rd("h = 8 (wide)",   rdrobust(df_muni[[V_OUTCOME]], df_muni$r, c = 0, h = 8, masspoints = "off")),
  tryCatch(collect_rd("Donut r = 1", rdrobust(filter(df_muni, abs(r) > 1)[[V_OUTCOME]],
                                               filter(df_muni, abs(r) > 1)$r, c = 0, h = 6, masspoints = "off")),
           error = function(e) data.frame(label = "Donut r = 1", est = NA, lo = NA, hi = NA)),
  tryCatch(collect_rd("Donut r = 2", rdrobust(filter(df_muni, abs(r) > 2)[[V_OUTCOME]],
                                               filter(df_muni, abs(r) > 2)$r, c = 0, h = 6, masspoints = "off")),
           error = function(e) data.frame(label = "Donut r = 2", est = NA, lo = NA, hi = NA))
) %>% mutate(label = factor(label, levels = rev(label)))

ggplot(specs, aes(x = est, y = label)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  geom_pointrange(aes(xmin = lo, xmax = hi), color = "#800000", size = 0.4) +
  labs(x = "RD estimate (total casualties)", y = NULL,
       title = "Specification chart: main result across six specs") +
  theme_minimal(base_size = 11) +
  theme(plot.title = element_text(color = "#800000"))
```

**[Write ~350 words here.]** Key points to address:

- Report the paper's published estimate and whether your replication hits it.
  If not, state what differs and why (variable map, aggregation level).
- Report the MSE-optimal estimate (Robust row). State the optimal bandwidth
  and whether it is close to the paper's h = 6.
- Interpret the donut results: do estimates grow, shrink, or hold stable as
  the hole widens? Relate back to the misclassification argument in Section 1.
- One sentence on bandwidth sensitivity: are the h/2 and 2h estimates inside
  each other's CIs?

---

## Section 3 — Timing and Actor Validity

*(~300 words, two committed verdicts | Stage 3: 10:40–11:10)*

```{r}
#| label: actor-splits

cat("=== Insurgent-initiated ===\n")
summary(rdrobust(df_muni[[V_CAS_INS]], df_muni$r, c = 0, h = 6, masspoints = "off"))

cat("\n=== Government-initiated ===\n")
summary(rdrobust(df_muni[[V_CAS_GOV]], df_muni$r, c = 0, h = 6, masspoints = "off"))
```

```{r}
#| label: measurement-probe

# Do the initiator splits sum to the total at the cutoff?
# A gap suggests initiator coding varies with program presence.
ins <- rdrobust(df_muni[[V_CAS_INS]], df_muni$r, c = 0, h = 6, masspoints = "off")$coef["Robust", 1]
gov <- rdrobust(df_muni[[V_CAS_GOV]], df_muni$r, c = 0, h = 6, masspoints = "off")$coef["Robust", 1]
tot <- rdrobust(df_muni[[V_OUTCOME]],  df_muni$r, c = 0, h = 6, masspoints = "off")$coef["Robust", 1]

cat(sprintf("Insurgent RD: %.3f\nGovernment RD: %.3f\nSum of splits: %.3f\n",
            ins, gov, ins + gov))
```

**[Write ~300 words here with two committed verdicts.]** Structure:

- **Verdict 1 — Actor mechanism.** Which initiator drives the total effect?
  The paper argues insurgents preemptively escalate violence to sabotage the
  program. Is the insurgent-initiated jump larger than the government-initiated
  jump? State a verdict: consistent / inconsistent with the preemption story.

- **Verdict 2 — Measurement.** Do the two actor splits sum to the total at
  the cutoff? A gap would suggest initiator attribution (who started the
  fight) is not independent of program presence — a coding validity problem.
  State a verdict: initiator coding appears / does not appear reliable.

- **The stage-window problem.** The paper pools all periods into a single RD.
  Under staggered rollout, municipalities received the program at different
  times. Pooling pre- and post-program periods conflates the intent-to-treat
  with a noisy average that includes pre-program windows where the effect
  should be zero. Section 4's RD event study is the direct answer.

---

## Section 4 — The Complement and a Better Clock

*(~300 words + Figure Array 2 | Stage 4: 11:10–11:45)*

```{r}
#| label: did-event-study
#| fig-width: 8
#| fig-height: 4.2
#| fig-cap: "DiD event study: eligible vs. ineligible municipalities around program start. Pre-trend coefficients near zero support parallel trends; post-period estimates show the dynamic effect."

df <- df %>%
  mutate(
    k     = .data[[V_TIME]] - .data[[V_START]],
    k_bin = pmax(pmin(k, 8), -8),
    k_es  = if_else(elig == 1, k_bin, -1L)
  )

m_es <- feols(
  as.formula(paste0(V_OUTCOME, " ~ i(k_es, ref = -1) | ", V_MUNI, " + ", V_TIME)),
  data    = df,
  cluster = as.formula(paste0("~", V_MUNI))
)

iplot(m_es,
      main   = "DiD event study: eligible vs. ineligible",
      xlab   = "event time (periods since program start)",
      ylab   = "effect on conflict casualties")
```

```{r}
#| label: rd-event-study
#| fig-width: 8
#| fig-height: 4.2
#| fig-cap: "RD event study: period-by-period jumps at the poverty-rank cutoff. Leads (k < 0) serve as built-in placebos. The path from onset to peak to fade replaces the paper's three coarse period splits."

rd_path <- lapply(sort(unique(df$k_bin)), function(kk) {
  d <- filter(df, k_bin == kk) %>%
    distinct(.data[[V_MUNI]], .keep_all = TRUE)
  if (nrow(d) < 10) return(NULL)
  fit <- tryCatch(rdrobust(d[[V_OUTCOME]], d$r, c = 0, h = 6, masspoints = "off"),
                  error = function(e) NULL)
  if (is.null(fit)) return(NULL)
  data.frame(k   = kk,
             est = fit$coef["Robust", 1],
             lo  = fit$ci["Robust", 1],
             hi  = fit$ci["Robust", 2])
}) %>% bind_rows()

if (nrow(rd_path) == 0) {
  cat("Note: no event-time periods had sufficient observations for rdrobust.\n",
      "Check V_START variable mapping — aff1 may not correctly identify program start timing.\n")
} else {
  print(
    ggplot(rd_path, aes(x = k, y = est)) +
      geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
      geom_vline(xintercept = -0.5, linetype = "dashed", color = "gray50") +
      geom_pointrange(aes(ymin = lo, ymax = hi), color = "#800000", size = 0.35) +
      labs(x = "event time (periods since program start)",
           y = "RD estimate (casualties)",
           title = "RD event study: period-by-period jumps at the cutoff") +
      theme_minimal(base_size = 11) +
      theme(plot.title = element_text(color = "#800000"))
  )
}
```

**[Write ~300 words here.]** Key points to address:

- **DiD pre-trends.** Are the pre-period coefficients (k < −1) flat and near
  zero? Interpret this as evidence for or against parallel trends between
  eligible and ineligible municipalities. Apply the full Tuesday checklist:
  reference period choice, endpoint bins, window length.

- **RD event study: onset/peak/fade.** Describe the shape of the path. When
  does the effect emerge, peak, and (if visible) decay? The leads (k < 0)
  are built-in placebos — if they are near zero, the identifying assumption
  is credible period by period.

- **Why the RD event study is a better clock.** The paper collapses dynamics
  into three coarse periods; the RD event study traces the full path. Wide
  intervals are expected (one period's data per estimate) — pair the dynamic
  picture with the pooled estimate: shape from the path, precision from the
  pooled RD.

- **Which design you trust more and why.** The DiD requires parallel trends
  across the eligibility boundary over time; the RD requires continuity at
  the rank cutoff within each period. State which assumption you find more
  credible given the data and the setting.
